Operational Feature Selection with Gaussian Mixture Models for remote sensing images

### 2016-10-12 paper.tex
add _c+change for crit evolution+add line in table 7

        \emph{The   Kullback-Leibler   divergence}   (KL   divergence)
        measures  the  amount  of  information  lost  when  the  first
        distribution     is     approximated     by     the     second
        one\cite{kullback1987letter}. It can be explicitly computed in
        the case of Gaussian distributions:
        \begin{align}
            &KL_{ij} = \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j) \nonumber \\
            & + (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \boldsymbol{\Sigma}_i^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - d + \log \left( \frac{|\boldsymbol{\Sigma}_i|}{|\boldsymbol{\Sigma}_j|} \right) \biggr\},
        \end{align}
        where Tr is the trace operator and $d$ the dimension of the distribution.

        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
        $KL_{ij} \ne KL_{ji}$.  A symmetrized version  is used to
        compute the criterion function:
        \begin{align}\label{eq:skl}
          SKL_{ij} &=KL_{ij} + KL_{i,j} \nonumber \\
            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_i^{-1} \boldsymbol{\Sigma}_j + \boldsymbol{\Sigma}_j^{-1} \boldsymbol{\Sigma}_i) \nonumber \\
            &~~+ (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t (\boldsymbol{\Sigma}_i^{-1} + \boldsymbol{\Sigma}_j^{-1}) (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) - 2d \biggr\}.
        \end{align}

        The extension  to the multiclass  situation is done  by taking
        the weighted mean  of the KL divergences computed  on all pair
        of classes~\cite{bruzzone1995extension}:
        \begin{equation}
            C_{SKL} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j SKL_{ij}.
        \end{equation}

        \emph{The Bhattacharyya distance} is defined in the case of Gaussian model as
        \begin{align}
            {B}_{ij} = &\frac{1}{8} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j)^t \left( \frac{\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j}{2} \right)^{-1} (\boldsymbol{\mu}_i - \boldsymbol{\mu}_j) \nonumber \\
            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j|}{\sqrt{|\boldsymbol{\Sigma}_i| |\boldsymbol{\Sigma}_j|}} \right).
        \end{align}

        The \emph{Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance. It saturates if the separability between the two distribution increases~\cite{bruzzone2009novel}. The JM distance is defined as
        \begin{equation}\label{eq:jm}
            {JM}_{ij} = \sqrt{ 2 \{1 - \text{exp}(-B_{ij})\} }.
        \end{equation}

        Similar to the KL divergence,  a weighted mean of the distance
        between two classes is computed to aggregate the measures in a
        single value:
        \begin{equation}
            {C}_{JM} = \sum_{i=1}^{C} \sum_{j=i + 1}^{C} \pi_i \pi_j {JM}_{ij}.
        \end{equation}

@@

        In the  remaining of the  paper, $\boldsymbol{\Sigma}^{(k-1)}$
        denotes the  covariance matrix of the  $(k-1)^{th}$ iteration,
        \emph{i.e.}, the  covariance matrix  of the  selected features
        and $\boldsymbol{\Sigma}^{(k)}$ denotes a covariance matrix at
        the $k^{th}$ iteration, \emph{i.e.},  the covariance matrix of
        an augmented set. Then, since $\boldsymbol{\Sigma}^{(k)}$ is a
        positive definite symmetric matrix,  the covariance matrix can
        be written as

@@

        \begin{equation}\label{eq:cov}
            \boldsymbol{\Sigma}^{(k)} =
            \bigg[\begin{array}{cc}
            \boldsymbol{\Sigma}^{(k-1)} & \mathbf{u}      \\
            \mathbf{u}^t          & \sigma_{kk} \\
            \end{array}\bigg],
        \end{equation}
        where  $\mathbf{u}$  is  the  $k^{th}$ column  of  the  matrix
        without      the      diagonal      element,      \emph{i.e.},
        $\mathbf{u}_{i}   =    \boldsymbol{\Sigma}^{(k)}_{i,k}$   with
        $i  \in [1,k-1]$.  Using  block matrix  inverse formulae,  the
        following proposition holds:
        \begin{prop}[Forward update rule for the inverse of the covariance matrix]
        \label{eq:update-inv}
        \begin{equation}\label{eq:cov:inv}
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                \mathbf{A} & \mathbf{v} \\
                \mathbf{v}^t  & \frac{1}{\alpha} \\
                \end{array}\bigg]
            \end{equation}
        \end{prop}
        \noindent                                                where
        $\mathbf{A}     =     (\boldsymbol{\Sigma}^{(k-1)})^{-1}     +
        \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}
        \mathbf{u}^t              (\boldsymbol{\Sigma}^{(k-1)})^{-1}$,
        $\mathbf{v}           =           -           \frac{1}{\alpha}
        (\boldsymbol{\Sigma}^{(k-1)})^{-1}       \mathbf{u}$       and
        $      \alpha      =      \sigma_{kk}      -      \mathbf{u}^t
        (\boldsymbol{\Sigma}^{(k-1)})^{-1}  \mathbf{u}$.  This  update
        formulae      is     used      to      obtain     all      the
        $(\boldsymbol{\Sigma}^{(k)})^{-1}$  corresponding  to all  the
        possible augmented set of a given selection iteration. Similar
        update formulae can be written for the backward step.

@@

        Using eq.~(\ref{eq:cov}) and~(\ref{eq:cov:inv}), it is possible to
        deduce  the   following  propositions  (proof  are   given  in
        Appendix~\ref{app:proof-update}).
        \begin{prop}[Update rule for the quadratical term]
        \label{eq:update-quad}
            \begin{align}
              (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)})^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once per selection step}}} \nonumber \\
                                                                                       &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for each augmented set}}}.
            \end{align}
        \end{prop}
        \begin{prop}[Update rule for logdet]
        \label{eq:update-log}
            \begin{equation}
                \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
            \end{equation}
        \end{prop}

@@

        \begin{prop}[Decision function~(\ref{eq:decision})]
            \begin{align}
            \label{eq:q-update}
                Q(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once per selection step}}} \nonumber \\
                &- \underbrace{\alpha \left( \left[\begin{array}{cc} \mathbf{v}^t & \frac{1}{\alpha} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}^{(k)}) \right)^2 - \log \alpha}_{\substack{\text{computed for each augmented set}}}.
            \end{align}
        \end{prop}

      \begin{prop}[Kullback-Leibler divergence~(\ref{eq:skl})]
        \begin{align}
        \label{eq:skl-update}
            & SKL_{ij} = \frac{1}{2} \biggl\{ \underbrace{\text{\normalfont Tr} \left((\boldsymbol{\Sigma}_i^{(k)})^{-1} \boldsymbol{\Sigma}_j^{(k)} + (\boldsymbol{\Sigma}_j^{(k)})^{-1} \boldsymbol{\Sigma}_i^{(k)}\right)}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_i}^t & \frac{1}{\alpha_i} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v_j}^t & \frac{1}{\alpha_j} \end{array}\right] (\boldsymbol{\mu}_i^{(k)} - \boldsymbol{\mu}_j^{(k)}) )^2 - 2k}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t (\boldsymbol{\Sigma}_i^{(k-1)})^{-1}(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
          & + \underbrace{(\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t(\boldsymbol{\Sigma}_j^{(k-1)})^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
        \end{align}
        with $(\boldsymbol{\Sigma}_i^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
      \end{prop}

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            \text{B}_{ij} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_i^{(k-1)} - \boldsymbol{\mu}_j^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_i^{(k-1)}| |\boldsymbol{\Sigma}_j^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_{i}^{(k)} - \boldsymbol{\mu}_{j}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_i \alpha_j}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
      where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_i + \boldsymbol{\Sigma}_j$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha$ and $\mathbf{v}$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}$.
    \end{prop}

@@

    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}$:

    \begin{eqnarray}
    \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} \boldsymbol{\Lambda}^{(k)} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp}
    \end{eqnarray}

    where  $\boldsymbol{\Lambda}^{(k)}$ and  $\mathbf{Q}^{(k)}$ are  the
    diagonal matrix  of eigenvalues of  the covariance matrix  and the
    orthonormal matrix of corresponding eigenvectors, respectively. To
    prevent numerical instability, the  eigenvalues are thresholded to
    EPS\_FLT. In our  implementation, EPS\_FLT is set  to the floating
    machine precision.

@@

    \begin{eqnarray}
      \boldsymbol{\Sigma}^{(k)} = \mathbf{Q}^{(k)} (\tilde{\boldsymbol{\Lambda}}^{(k)})^{-1} (\mathbf{Q}^{(k)})^t\label{eq:eigendecomp:inv}
    \end{eqnarray}

@@

    \begin{eqnarray}
      \label{eq:log:det}
      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_i^{(k)}),
    \end{eqnarray}
    where the \textasciitilde{} indicated thresholded values.

@@

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
            \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
                \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
                \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
            \end{axis};
            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_50000spl.txt};
            %     \draw[thick,red] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            %     \draw[thick,black] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            % \end{axis};
        \end{tikzpicture}
        \caption{Criterion evolution in function of the number of selected variables. Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
    \end{figure}

@@

    \begin{table*}[!t]
        \centering
        \caption{Kappa coefficient and processing time for 50,000 samples by class and averaged over 5 trials (standard deviation in parenthesis). Processing times are given in second.\label{tab:potsdam-otbsimu-big}}
        \begin{tabular}{lcccccc}\toprule
            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cmidrule{2-7}
            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\
            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6 & 330 & 10 \\
            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6.6 & 340 & 10 \\
            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 0.440 (0.015) & 460 & 2000 & all \\
            Random Forest & {\bfseries 0.851 (0.001)} & {\bfseries 0.715 (0.001)} & {\bfseries 0.573 (0.002)} & 2000 & 2000 & all \\
            \bottomrule
        \end{tabular}
    \end{table*}

>>>>

        \emph{The   Kullback-Leibler   divergence}   (KL   divergence)
        measures  the  amount  of  information  lost  when  the  first
        distribution     is     approximated     by     the     second
        one\cite{kullback1987letter}. It can be explicitly computed in
        the case of Gaussian distributions:
        \begin{align}
            &KL_{cc^\prime} = \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_c^{-1} \boldsymbol{\Sigma}_{c^\prime}) \nonumber \\
            & + (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \boldsymbol{\Sigma}_c^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) - d + \log \left( \frac{|\boldsymbol{\Sigma}_c|}{|\boldsymbol{\Sigma}_{c^\prime}|} \right) \biggr\},
        \end{align}
        where Tr is the trace operator and $d$ the dimension of the distribution.

        The    KL   divergence    is   not    symmetric,   \emph{i.e.}
        $KL_{cc^\prime} \ne KL_{c^\prime c}$.  A symmetrized version  is used to
        compute the criterion function:
        \begin{align}\label{eq:skl}
          SKL_{cc^\prime} &=KL_{cc^\prime} + KL_{c^\prime c} \nonumber \\
            &= \frac{1}{2} \biggl\{ \text{Tr} (\boldsymbol{\Sigma}_c^{-1} \boldsymbol{\Sigma}_{c^\prime} + \boldsymbol{\Sigma}_{c^\prime}^{-1} \boldsymbol{\Sigma}_c) \nonumber \\
            &~~+ (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t (\boldsymbol{\Sigma}_c^{-1} + \boldsymbol{\Sigma}_{c^\prime}^{-1}) (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) - 2d \biggr\}.
        \end{align}

        The extension  to the multiclass  situation is done  by taking
        the weighted mean  of the KL divergences computed  on all pair
        of classes~\cite{bruzzone1995extension}:
        \begin{equation}
            C_{SKL} = \sum_{c=1}^{C} \sum_{c^\prime = c + 1}^{C} \pi_c \pi_{c^\prime} SKL_{cc^\prime}.
        \end{equation}

        \emph{The Bhattacharyya distance} is defined in the case of Gaussian model as
        \begin{align}
            {B}_{cc^\prime} = &\frac{1}{8} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime})^t \left( \frac{\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}}{2} \right)^{-1} (\boldsymbol{\mu}_c - \boldsymbol{\mu}_{c^\prime}) \nonumber \\
            &+ \frac{1}{2} \ln \left( \frac{|\boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}|}{\sqrt{|\boldsymbol{\Sigma}_c| |\boldsymbol{\Sigma}_{c^\prime}|}} \right).
        \end{align}

        The \emph{Jeffries–Matusita distance} is a measure based on the Bhattacharyya distance. It saturates if the separability between the two distribution increases~\cite{bruzzone2009novel}. The JM distance is defined as
        \begin{equation}\label{eq:jm}
            {JM}_{cc^\prime} = \sqrt{ 2 \{1 - \text{exp}(-B_{cc^\prime})\} }.
        \end{equation}

        Similar to the KL divergence,  a weighted mean of the distance
        between two classes is computed to aggregate the measures in a
        single value:
        \begin{equation}
            {C}_{JM} = \sum_{c=1}^{C} \sum_{c^{\prime} = c + 1}^{C} \pi_c \pi_{c^\prime} {JM}_{cc^\prime}.
        \end{equation}

@@

        In the  remaining of the  paper, $\boldsymbol{\Sigma}_c^{(k-1)}$
        denotes the  covariance matrix of the  $(k-1)^{th}$ iteration,
        \emph{i.e.}, the  covariance matrix  of the  selected features
        and $\boldsymbol{\Sigma}_c^{(k)}$ denotes a covariance matrix at
        the $k^{th}$ iteration, \emph{i.e.},  the covariance matrix of
        an augmented set. Then, since $\boldsymbol{\Sigma}_c^{(k)}$ is a
        positive definite symmetric matrix,  the covariance matrix can
        be written as

@@

        \begin{equation}\label{eq:cov}
            \boldsymbol{\Sigma}^{(k)} =
            \bigg[\begin{array}{cc}
            \boldsymbol{\Sigma}^{(k-1)}_c & \mathbf{u}_c      \\
            \mathbf{u}_c^t          & \sigma^{(k)}_c \\
            \end{array}\bigg],
        \end{equation}
        where  $\mathbf{u}_c$  is  the  $k^{th}$ column  of  the  matrix
        without      the      diagonal      element,      \emph{i.e.},
        $\mathbf{u}_{c}(i)   =    \boldsymbol{\Sigma}^{(k)}_{c}(i,k)$   with
        $i  \in [1,k-1]$.  Using  block matrix  inverse formulae,  the
        following proposition holds:
        \begin{prop}[Forward update rule for the inverse of the covariance matrix]
        \label{eq:update-inv}
        \begin{equation}\label{eq:cov:inv}
                (\boldsymbol{\Sigma}^{(k)})^{-1} =
                \bigg[\begin{array}{cc}
                \mathbf{A}_c & \mathbf{v}_c \\
                \mathbf{v}_c^t  & \frac{1}{\alpha_c} \\
                \end{array}\bigg]
            \end{equation}
        \end{prop}
        \noindent                                                where
        $\mathbf{A}_c     =     (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}     +
        \frac{1}{\alpha_c} (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} \mathbf{u}_c
        \mathbf{u}_c^t              (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}$,
        $\mathbf{v}_c           =           -           \frac{1}{\alpha_c}
        (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}       \mathbf{u}_c$       and
        $      \alpha_c      =      \sigma^{(k)}_c      -      \mathbf{u}_c^t
        (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1}  \mathbf{u}_c$.  This  update
        formulae      is     used      to      obtain     all      the
        $(\boldsymbol{\Sigma}^{(k)}_c)^{-1}$  corresponding  to all  the
        possible augmented set of a given selection iteration. Similar
        update formulae can be written for the backward step.

@@

        Using eq.~(\ref{eq:cov}) and~(\ref{eq:cov:inv}), it is possible to
        deduce  the   following  propositions  (proof  are   given  in
        Appendix~\ref{app:proof-update}).
        \begin{prop}[Update rule for the quadratical term]
        \label{eq:update-quad}
            \begin{align}
              (\mathbf{x}^{(k)})^t (\boldsymbol{\Sigma}^{(k)}_c)^{-1} \mathbf{x}^{(k)} = &\underbrace{(\mathbf{x}^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} \mathbf{x}^{(k-1)}}_{\substack{\text{computed once per selection step}}} \nonumber \\
                                                                                       &+ \underbrace{\alpha_c ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] \mathbf{x}^{(k)} )^2}_{\substack{\text{computed for each augmented set}}}.
            \end{align}
        \end{prop}
        \begin{prop}[Update rule for logdet]
        \label{eq:update-log}
            \begin{equation}
                \log \left(|\boldsymbol{\Sigma}^{(k)}_c|\right) = \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right)}_{\substack{\text{computed once}\\ \text{per selection step}}} + \underbrace{\log \alpha_c}_{\substack{\text{computed for} \\ \text{each augmented set}}}.
            \end{equation}
        \end{prop}

@@

        \begin{prop}[Decision function~(\ref{eq:decision})]
            \begin{align}
            \label{eq:q-update}
                Q_c(\mathbf{x}) = &- \underbrace{(\mathbf{x}^{(k-1)} - \boldsymbol{\mu}_c^{(k-1)})^t (\boldsymbol{\Sigma}^{(k-1)}_c)^{-1} (\mathbf{x}^{(k-1)} - \boldsymbol{\mu}^{(k-1)}_c)}_{\substack{\text{computed once per selection step}}} \nonumber \\
                &- \underbrace{\log \left(|\boldsymbol{\Sigma}^{(k-1)}_c|\right) + 2 \log (\pi_c) + k \log(2 \pi)}_{\substack{\text{computed once per selection step}}} \nonumber \\
                &- \underbrace{\alpha_c \left( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] (\mathbf{x}^{(k)} - \boldsymbol{\mu}_c^{(k)}) \right)^2 - \log \alpha_c}_{\substack{\text{computed for each augmented set}}}.
            \end{align}
        \end{prop}

      \begin{prop}[Kullback-Leibler divergence~(\ref{eq:skl})]
        \begin{align}
        \label{eq:skl-update}
            & SKL_{cc^\prime} = \frac{1}{2} \biggl\{ \underbrace{\text{\normalfont Tr} \left((\boldsymbol{\Sigma}_c^{(k)})^{-1} \boldsymbol{\Sigma}_{c^\prime}^{(k)} + (\boldsymbol{\Sigma}_{c^\prime}^{(k)})^{-1} \boldsymbol{\Sigma}_c^{(k)}\right)}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}_c^t & \frac{1}{\alpha_c} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{\alpha ( \left[\begin{array}{cc} \mathbf{v}_{c^\prime}^t & \frac{1}{\alpha_{c^\prime}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 - 2k}_{\substack{\text{computed for each augmented set}}} \nonumber \\
            &+ \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber\\
          & + \underbrace{(\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t(\boldsymbol{\Sigma}_{c^\prime}^{(k-1)})^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \biggr\}.
        \end{align}
        with $(\boldsymbol{\Sigma}_c^{(k)})^{-1}$ computed with Proposition~\ref{eq:update-inv};
      \end{prop}

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \ln \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \ln \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
    \end{prop}
    \noindent where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha_c$ and $\mathbf{v}_c$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}_c$.

@@

    To deal with this limitation, the choice has been made to perform an eigendecomposition of the covariance matrix $\boldsymbol{\Sigma}^{(k)}_c$:

    \begin{eqnarray}
    \boldsymbol{\Sigma}^{(k)}_c = \mathbf{P}^{(k)}_c \boldsymbol{\Lambda}^{(k)}_c (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp}
    \end{eqnarray}

    where  $\boldsymbol{\Lambda}^{(k)}_c$ and  $\mathbf{P}^{(k)}_c$ are  the
    diagonal matrix  of eigenvalues of  the covariance matrix  and the
    orthonormal matrix of corresponding eigenvectors, respectively. To
    prevent numerical instability, the  eigenvalues are thresholded to
    EPS\_FLT. In our  implementation, EPS\_FLT is set  to the floating
    machine precision.

@@

    \begin{eqnarray}
      (\boldsymbol{\Sigma}^{(k)}_c)^{-1} = \mathbf{P}^{(k)}_c (\tilde{\boldsymbol{\Lambda}}^{(k)}_c)^{-1} (\mathbf{P}^{(k)}_c)^t\label{eq:eigendecomp:inv}
    \end{eqnarray}

@@

    \begin{eqnarray}
      \label{eq:log:det}
      \log \left(|\boldsymbol{\Sigma}^{(k)}|\right) = \sum_{i=1}^d\log(\tilde{\lambda}_{c}^{(k)}(i)),
    \end{eqnarray}
    where the \textasciitilde{} indicated thresholded values and $\lambda_{c}(i)$ the ith eigenvalue.

@@

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
            \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl_aisa.txt};
                % \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
                % \draw[thick,black] (axis cs:30,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:30,\pgfkeysvalueof{/pgfplots/ymax});
            \end{axis};
            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
            %     \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
            %     \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
            % \end{axis};
            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_50000spl.txt};
            %     \draw[thick,red] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            %     \draw[thick,black] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            % \end{axis};
        \end{tikzpicture}
        \caption{Criterion evolution in function of the number of selected variables for first trial with Aisa dataset. Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
    \end{figure}

@@

    \begin{table*}[!t]
        \centering
        \caption{Kappa coefficient and processing time for 50,000 samples by class and averaged over 5 trials (standard deviation in parenthesis). Processing times are given in second. NB: the test has not been conduct with KNN because of a too long processing time for classification.\label{tab:potsdam-otbsimu-big}}
        \begin{tabular}{lcccccc}\toprule
            & {\bfseries 5\_11 (train)} & {\bfseries 5\_12 (test)} & {\bfseries 3\_10 (test)} & {\bfseries Train. Time} & {\bfseries Classif. time} & {\bfseries \# of selected features} \\ \cmidrule{2-7}
            GMM SFS kappa & 0.713 (0.001) & 0.684 (0.001) & 0.531 (0.005) & 20000 & 340 & 29 \\
            GMM SFS JM &    0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6 & 330 & 10 \\
            GMM SFFS JM &   0.560 (0.111) & 0.576 (0.104) & 0.435 (0.085) & 6.6 & 340 & 10 \\
            GMM ridge &     0.641 (0.015) & 0.611 (0.026) & 0.440 (0.015) & 460 & 2000 & all \\
            KNN &           /             & /             & /             & / & / & / \\
            Random Forest & {\bfseries 0.851 (0.001)} & {\bfseries 0.715 (0.001)} & {\bfseries 0.573 (0.002)} & 2000 & 2000 & all \\
            \bottomrule
        \end{tabular}
    \end{table*}
